Handbook streamFind: backend

### 2022-07-23 handbooks/Handbook_backend_streamFind.Rmd
LFU09 and changed ms file names

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.align = "center", cache = TRUE) # fig.width = 9,
```
<br> <br>

@@

@@

@@

@@

```{r description-files, echo=FALSE}
df_files_desc <- data.frame(
  file_name = basename(files),
  device = c(rep("Agilent Q-TOF", 24)),
  type = c(rep("centroided", 3), rep("profile", 3), rep("profile", 18)),
  mode = c(rep("MS/MS", 24)),
  polarity = c(rep("positive", 6),
               rep("negative", 3), rep("positive",3),
               rep("negative", 3), rep("positive",3),
               rep("negative", 3), rep("positive",3)),
  description = c(
    rep("Basic centroided MS data spiked with chemicals.", 3),
    rep("Basic profile MS data spiked with chemicals.", 3),
    rep("Blank MS data spiked with internal standards.", 6),
    rep("Wastewater secondary effluent MS data spiked with internal standards.", 6),
    rep("Wastewater secondary effluent treated with ozonated strong water MS data spiked with internal standards.", 6)
  )
)

knitr::kable(df_files_desc, caption = "Files included in the streamFind package.") %>%
kable_styling(font_size = 11, bootstrap_options = c("striped", "hover", "condensed", "responsive"), fixed_thead = TRUE)
```

@@

```{r show-table-chemicals, echo=FALSE}
db$ionization <- "positive"
db[, `:=`("mz_pos" = neutralMass + 1.0073, "mz_neg" = neutralMass - 1.0073 )]
db[grepl("neg", db$comment), ionization := "both"]
data.table::setnames(db, "mixIUTA", "tag")
db[tag %in% 1, tag := "S"]
knitr::kable(db[, .(name, PubChem, formula, neutralMass, rt, ionization, tag)], caption = "Chemicals spiked.") %>%
kable_styling(font_size = 11, bootstrap_options = c("striped", "hover", "condensed", "responsive"), fixed_thead = TRUE)
```

@@

```{r classes-overview, echo=FALSE, fig.cap="Overview of the S4 classes and their hierarchy with emphasis on MS analyses.", out.width="95%"}
knitr::include_graphics(paste0(getwd(), "/hb_backend_figures/classes_overview.png"))
```

@@

* _streamProject_ refers to the broader S4 class applied in the streamFind package and holds basic project information, see \@ref(streamproject-class);
* _msData_ refers to the S4 subclass of _streamProject_ when MS analyses (i.e., MS files) are added to the project, see \@ref(msdata-class);
* _msAnalysis_ refers to the S4 class of the object generated for each MS file added to the _msData_ object in the slot "analyses";
* `projectInfo` is the method to obtain the basic information (i.e., title, date and path) of a _streamProject_ object;
* `analyses` is the name of the slot holding the analyses (as list) and is also a method to get the analysis names in the _streamProject_;
* `replicate` is the name of an analysis replicate group and the method to get the replicate names;
* `blank` is the name of a blank analysis replicate group to be used for blank (blind) subtraction and the method to get blank replicate names;
* __traces__ corresponds to a raw data signal (e.g., mass trace in MS data) obtained from a given analysis file;  
* `spectra` for MS analyses is the method to get a table with a spectrum for each cycle time (i.e., retention time
`r # TODO add chromatograms entry and make method`
* __EIC__ is an extracted ion chromatogram (retention time vs intensity) and can be obtained with the method `EICs`;
* __TIC__ is a total ion chromatograms (retention time vs intensity) and can be obtained with the method `TICs`;
* __XIC__ is a three dimensional ion chromatogram (_m/z_ vs retention time vs intensity) and can be obtained with the method `XICs`;
* __peak__ corresponds to a cluster of traces (e.g., integrated chromatographic MS peak) within an analysis file and can be accessed with the method `peaks`;
* __feature__ corresponds to a group of corresponding peaks across analysis files/analyses and can be accessed with the method `features`;
* __MS2__ corresponds to second order fragmentation data as acquired in tandem MS or MS/MS mode and can be accessed with the method `MS2s`.

(...)

@@

```{r newStreamProject-call, message=FALSE, results='hide'}
sp1 <- newStreamProject(files = files[1:3], title = "Example_01")
```

@@

```{r msData_streamProject_classes}
is(sp1)
```

The _msData_ structure includes the _streamProject_ class plus the `features` slot. The project, date and path slots correspond to the project title, creation date and directory in disk, respectively. The analyses slot is a list with a _msAnalysis_ S4 class object per MS file added. The structure of the _msAnalysis_ class is described in the next sub-chapter. The `features` slot is a _msFeatures_ S4 class object to hold the results from alignment, grouping and annotation of peaks across analyses.

@@

```{r structure_msData}
str(sp1, max.level = 2)
```

@@

The _msAnalysis_ class is structured as shown below. Note that the S4 method `getAnalyses` for _msData_ was used to obtain the _msAnalysis_ of the first analysis as defined by the second argument. In general, access and plotting S4 methods can also be applied directly to the _msAnalysis_ class besides _msData_ (as described in \@ref(low-level-data-access-and-plotting)) but is not part of the basic workflow and mostly used for advanced data visualization and access of low level data (e.g., traces).

@@

```{r msAnalysis_structure}
a1 <- getAnalyses(sp1, 1)
str(a1, max.level = 2)
```

@@

```{r EICs_method}
#EIC of targets from first analysis using msAnalysis class
eic1 <- EICs(a1, mz = targets4)
head(eic1, 6)

@@

```{r plotEICs_static}
#plot the produced EIC data.table
plotEICs(eic1)
```

```{r plotEICs_interactive}
#using the msData class object and targets but with interactive plot
plotEICs(sp1, analyses = NULL, mz = targets4, colorBy = "analyses", interactive = TRUE)

@@

```{r TICs_method}
# TIC for both analyses in the msData class object
tic1 <- TICs(sp1)
tic1

@@

```{r plotTICs_method}
#using a data.table as obatined by TICs
plotTICs(tic1, interactive = TRUE)

@@

```{r XICs_method}
#using the msData class object
xic1 <- XICs(sp1, mz = targets4)
xic1

@@

```{r plotXICs_method}
#using the output of XICs S4 method (a data.table)
plotXICs(xic1, legendNames = c("target number 1", "target number 2"), plotTargetMark = TRUE,
 targetsMark = targets4[, c("mz", "rt")], ppmMark = 5, secMark = 10, numberRows = 2)

@@

```{r loadRawData}
#load raw data to an msAnalysis object
a1 <- loadRawData(a1)

@@

```{r newStreamProject_basic_workflow, message=FALSE, results='hide'}
sp2 <- newStreamProject(files = files[c(10:12, 16:18)], path = getwd(), title = "Project Example")
```

@@

```{r assign_replicates}
#getter for replicate names in msData
replicates(sp2)

@@

```{r assign_blanks}
#getter for blank replicates in each sample in a msData object
blanks(sp2)

@@

```{r other_S4_methods}
#getter for analysis names in msdata
analyses(sp2)

@@

```{r create_processing_parameters}

sp2_pp <- sp2

@@

```{r peakPicking}

#peak picking call using the stored parameters
sp2_pp <- peakPicking(sp2_pp)

# peaks added for each analysis, as shown by column peaks of show method (below)
sp2_pp

#alternatively, parameters can be added during the function call
sp2_pp_2 <- peakPicking(sp2, settings = settings_pp)

# parameters are added to each analysis after running peak picking
summary(getParameters(sp2_pp_2, where = "analyses", analyses = 1))

sp2_pp_2

# running peak picking for an msAnalysis object (settings were already added above)
a2 <- peakPicking(a2)

# msAnalysis with peaks
a2
```

@@

```{r peakPicking_multiple_settings}

sp2_pp_3 <- sp2

@@

# the algorithm of the influent samples is different
sapply(analyses(sp2_pp_3), function(x) getAlgorithm(getParameters(sp2_pp_3, where = "analyses", analyses = x)[[1]]))

#running peak picking
sp2_pp_3 <- peakPicking(sp2_pp_3)

# as the peak parameters with "openms" as algorithm were less stringent, the influent samples have more peaks then when "xcms3" is used 
sp2_pp_3

@@

```{r export_import_settings}

@@

#export settings as JSON
exportSettings(settings_pp_2, name = "settings_pp_2", format = "json", path = getwd())

#import settings as rds
settings_pp_imported <- importSettings(file = paste0(getwd(), "/settings_pp.rds"))
all.equal(settings_pp, settings_pp_imported)

@@

```{r inspecting_peaks}
#access to peaks in msData based on built targets as used for extracting EICs
peaks(sp2_pp[c(1, 4)], mz = mz1[1], rt = rt1[1], ppm = 10, sec = 30)

@@

```{r creating_grouping_parameters}
param_g <- list(
  rtalign = TRUE,
  loadRawData = TRUE,
  groupParam = xcms::PeakDensityParam(
    sampleGroups = "holder",
    bw = 3,
    minFraction = 0.6,
    minSamples = 2,
    binSize = 0.008,
    maxFeatures = 100),
  preGroupParam = xcms::PeakDensityParam(
    sampleGroups = "holder",
    bw = 5,
    minFraction = 1,
    minSamples = 3,
    binSize = 0.008,
    maxFeatures = 100),
  retAlignParam = xcms::PeakGroupsParam(
    minFraction = 1,
    extraPeaks = 0,
    smooth = "loess",
    span = 0.3,
    family = "gaussian")
)

@@

```{r processing_grouping, message=FALSE}
#using the added parameters
sp2_pg <- peakGrouping(sp2_pp)

#adding the parameters as arguments
sp2_pg_2 <- peakGrouping(sp2_pp, settings = settings_pg)
```

@@

```{r msFeatures_class}
#accessing directly the msFeatures slot, not recommended
str(sp2_pg@features, max.level = 2)
```

@@

```{r inspecting_features}
#getter for feature intensities and intensity deviations
features(sp2_pg, mz = targets4)

@@

```{r plot_alignment}
plotAlignment(sp2_pg)
```

@@

```{r recursive_integration_settings}
#default settings S4 class with ChromPeakAreaParam from xcms as parameter settings
settings_fl <- fillingSettingsDefaultXCMS()

@@

```{r recursive_integration_procesing, results='hide', message=FALSE}
#filling peaks for msData
sp2_pg_fl <- peakFilling(sp2_pg, settings = settings_fl)
```

@@

```{r recursive_integration_inspection}
# inspecting the filled peaks for a given mz and rt target
plotFeaturePeaks(sp2_pg_fl, mz = 441.1670, rt = 916, ppm = 20, sec = 60)
```

@@

```{r annotation_processing, results='hide', message=FALSE}
#creating settings for using CAMERA
settings_an <- createSettings(
  call = "peakAnnotation",
  algorithm = "camera",
  settings = list(
    ionization = NA_character_,
    onlyIsotopes = FALSE,
    minSize = 1,
    relMinReplicates = 1,
    extraOpts = list(
      sigma = 6,
      perfwhm = 0.35,
      cor_eic_th = 0.3,
      graphMethod = "hcs",
      pval = 0.05,
      calcCiS = TRUE,
      calcIso = TRUE,
      calcCaS = TRUE,
      maxcharge = 3,
      maxiso = 5,
      ppm = 15,
      mzabs = 0.008,
      rules =  data.table::fread(system.file("rules/primary_adducts_pos.csv", package = "CAMERA"), header = TRUE),
      multiplier = 3,
      max_peaks = 500,
      intval = "maxo"
    )
  )
)

sp2_pg_fl_an <- peakAnnotation(sp2_pg_fl, settings = settings_an)
```

@@

```{r annotation_results}
#isotopic and adduct features annotated with a given target mz and rt
t(streamFind::annotation(sp2_pg_fl_an, mz = 441.1670, ppm = 10, rt = 916, sec = 5, all = FALSE))

#plotting the annotation of each feature in a given target
plotAnnotation(sp2_pg_fl_an, mz = 441.1670, ppm = 10, rt = 916, sec = 5, all = FALSE)
```

@@

```{r filter_subsetting}
#sub-setting the msData class on analyses (the 3 first)
sp2_pg[1:3, ]

#sub-setting the msData class on analys2s and features (the 3 last analyses and features annotated with a target)
target_s <- streamFind::annotation(sp2_pg_fl_an, mz = 441.1670, ppm = 10, rt = 916, sec = 5, all = FALSE)
sp2_pg_fl_an[4:6, target_s$id]

@@

```{r filter_features}

@@

```{r filter_peaks}

@@

```{r quality_sn}

@@

```{r create_settings_msn}
#the parameter settings to obtained and process MSn spectra
param_msn <- list(
  isolationTimeWindow = 0,
  isolationMassWindow = 1.3,
  clusteringMethod = "distance",
  clusteringUnit = "ppm",
  clusteringWindow = 10,
  minIntensityPre = 200,
  minIntensityPost = 300,
  asPatRoon = FALSE, #when TRUE, is only used by running generateMS2
  mergeVoltages = TRUE,
  mergeBy = "analyses"
)

@@

```{r extract_msn}
#head of MS2 data for target 1 in a msAnalysis object
head(MS2s(a1, mz = targets4[1, ], settings = settings_msn))

@@

@@

```{r demoCode, echo=FALSE, eval=FALSE, include=FALSE}

>>>>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.align = "center", fig.width = 9, cache = TRUE)
```
<br> <br>

@@

For installation of the streamFind package, it is recommended to firstly ensure that the dependencies are currently installed. Several existing R packages and external functionalities are used in streamFind for various processing steps. A major dependency of streamFind is the [patRoon](https://github.com/rickhelmus/patRoon) R package.`r # TODO add reference` The patRoon package combines several functionalities for basic and advanced data processing and can be used interchangeably with streamFind, see \@ref(patroon-interchangeability) for more information. Most of the tools that patRoon depends on are equally used in streamFind. In the next sub-chapter, the mandatory and optional dependencies are described with installation instructions. For a standalone and self-contained usage option, the streamFind docker image is recommended (see \@ref(docker-image)), avoiding the need to installing all required dependencies.

@@

### R and RTools

[R](https://cran.r-project.org/) and [RTools](https://cran.r-project.org/bin/windows/Rtools/) are essential dependencies. The `r R.version.string` can be obtained in https://cran.r-project.org/. The RTools can be downloaded in https://cran.r-project.org/bin/windows/Rtools/. Make sure to download the right version for the R version installed. Installation instructions are given in both sources.

### Third party software

The external third party dependencies recommended are listed in the section [other dependencies](https://rickhelmus.github.io/patRoon/handbook_bd/manual-installation.html#other-dependencies) from the patRoon [handbook](https://rickhelmus.github.io/patRoon/handbook_bd/manual-installation.html#r-prerequisites). Below we list the mandatory and really recommended software and resources.

* [Java JDK](https://www.oracle.com/java/technologies/downloads/)  
* [ProteoWizard](https://proteowizard.sourceforge.io/)  
* [OpenMS](https://www.openms.de/downloads/)  
* [MetFrag CL](http://ipb-halle.github.io/MetFrag/projects/metfragcl/)
* [PubChemLite](https://zenodo.org/record/6503754#.YtqxE3bP1hE)
* [SIRIUS](https://bio.informatik.uni-jena.de/software/sirius/)  
* [BioTransformer](https://bitbucket.org/djoumbou/biotransformer/downloads/)  
* [SAFD](https://bitbucket.org/SSamanipour/safd.jl/src/master/)  
* [Open Babel](http://openbabel.org/wiki/Main_Page)

### R packages

The following code lines are adapted from the patRoon [handbook](https://rickhelmus.github.io/patRoon/handbook_bd/manual-installation.html#r-prerequisites) to ensure that both patRoon and other mutual R dependencies are installed properly.

```{r r-packages, eval=FALSE}
install.packages(c("BiocManager", "remotes"))

BiocManager::install(c("mzR", "xcms","CAMERA"))

remotes::install_github("blosloos/enviPick")

BiocManager::install("ropls")
remotes::install_github("rickhelmus/KPIC2")

BiocManager::install("InterpretMSSpectrum")
remotes::install_github("cbroeckl/RAMClustR")

remotes::install_github("blosloos/nontargetData")
remotes::install_github("blosloos/nontarget")

remotes::install_github("rickhelmus/cliqueMS")

# only needed for Bruker DataAnalysis integration
remotes::install_github("BSchamberger/RDCOMClient")

BiocManager::install(c("BiocStyle", "Rgraphviz")) 
remotes::install_github("KelseyChetnik/MetaClean")

remotes::install_github("rickhelmus/patRoon", upgrade = "never", dependencies = TRUE)
```

After installation of the patRoon R package, the location of each third party software should be given in an `.Rprofile` file as suggested in section [other dependencies](https://rickhelmus.github.io/patRoon/handbook_bd/manual-installation.html#other-dependencies) the patRoon [handbook](https://rickhelmus.github.io/patRoon/handbook_bd/manual-installation.html#r-prerequisites). For facilitation the same code lines are given below. The function `patRoon::verifyDependencies()` from patRoon can be used to verify if all dependencies are found.

```{r options-third-party-software, eval=FALSE}
options(patRoon.path.pwiz = "C:/ProteoWizard") # location of ProteoWizard installation folder
options(patRoon.path.SIRIUS = "C:/sirius-win64-3.5.1") # location where SIRIUS was extracted
options(patRoon.path.OpenMS = "/usr/local/bin") # directory with the OpenMS binaries
options(patRoon.path.pngquant = "~/pngquant") # directory containing pngquant binary
options(patRoon.path.MetFragCL = "~/MetFragCommandLine-2.4.8.jar") # full location to the jar file
options(patRoon.path.MetFragCompTox = "C:/CompTox_17March2019_SelectMetaData.csv") # full location to desired CompTox CSV file
options(patRoon.path.MetFragPubChemLite = "~/PubChemLite_exposomics_20220429.csv") # full location to desired PubChemLite CSV file
options(patRoon.path.BioTransformer = "~/biotransformer/biotransformer-3.0.1.jar")
options(patRoon.path.obabel = "C:/Program Files/OpenBabel-3.0.0") # directory with Open Babel binaries
```

@@

The streamFind package can be installed from the GitHub repository https://github.com/ricardobachertdacunha/streamFind with the following code line.

```{r install-streamfind, eval=FALSE}
remotes::install_github("ricardobachertdacunha/streamFind", dependencies = TRUE)
```

## Docker image

@@

```{r description-files, echo=FALSE}
df_files_desc <- data.frame(
  N. = seq_len(length(files)),
  file_name = basename(files),
  device = c(rep("Agilent Q-TOF", 24)),
  type = c(rep("centroided", 3), rep("profile", 3), rep("profile", 18)),
  mode = c(rep("MS/MS", 24)),
  polarity = c(rep("positive", 6),
               rep("negative", 3), rep("positive",3),
               rep("negative", 3), rep("positive",3),
               rep("negative", 3), rep("positive",3)),
  description = c(
    rep("Basic centroided MS data spiked with chemicals.", 3),
    rep("Basic profile MS data spiked with chemicals.", 3),
    rep("Blank MS data spiked with internal standards.", 6),
    rep("Wastewater secondary effluent MS data spiked with internal standards.", 6),
    rep("Wastewater secondary effluent treated with ozonated strong water MS data spiked with internal standards.", 6)
  )
)

knitr::kable(df_files_desc, caption = "Files included in the streamFind package.") %>%
  column_spec(column = 2, width = "100px") %>%
  column_spec(column = 3, width = "60px") %>%
  kable_styling(font_size = 11.5, bootstrap_options = c("striped", "hover", "responsive"), fixed_thead = TRUE) %>% #"condensed",
  scroll_box(width = "100%", height = "600px")
```

@@

```{r show-table-chemicals, echo=FALSE}
db$ionization <- "positive"
db[, `:=`("mz_pos" = neutralMass + 1.0073, "mz_neg" = neutralMass - 1.0073 )]
db[grepl("neg", db$comment), ionization := "both"]
data.table::setnames(db, "mixIUTA", "tag")
db[tag %in% 1, tag := "S"]

knitr::kable(db[, .(name, CAS, formula, neutralMass, rt, ionization, tag)], caption = "Chemicals spiked.") %>%
  kable_styling(font_size = 11, bootstrap_options = c("striped", "hover", "responsive"), fixed_thead = TRUE) %>% #"condensed",
  scroll_box(width = "100%", height = "600px")
```

@@

```{r classes-overview, echo=FALSE, fig.cap="Overview of the S4 classes and their hierarchy with emphasis on MS analyses.", out.width="100%"}
knitr::include_graphics(paste0(getwd(), "/hb_backend_figures/classes_overview.png"))
```

@@

* _streamProject_ refers to the broader S4 class applied in the streamFind package and holds basic project information, see \@ref(streamproject-class);
* _msData_ refers to the S4 subclass of _streamProject_ when MS analyses (i.e., MS files) are added to the project, see \@ref(msdata-class);
* _msAnalysis_ refers to the S4 class of the object generated for each MS file added to the _msData_ object in the slot "analyses";
* `projectInfo` is the method to obtain the basic information (i.e., title, date and path) of a _streamProject_ object;
* `analyses` is the name of the slot holding the analyses (as list) and is also a method to get the analysis names in the _streamProject_;
* `replicate` is the name of an analysis replicate group and the method to get the replicate names;
* `blank` is the name of a blank analysis replicate group to be used for blank (blind) subtraction and the method to get blank replicate names;
* __traces__ corresponds to a raw data signal (e.g., mass trace in MS data) obtained from a given analysis file;
* `spectra` for MS analyses is the method to get a table with a spectrum for each cycle time (i.e., retention time);
* __EIC__ is an extracted ion chromatogram (retention time vs intensity) and can be obtained with the method `EICs`;
* __TIC__ is a total ion chromatograms (retention time vs intensity) and can be obtained with the method `TICs`;
* __XIC__ is a three dimensional ion chromatogram (_m/z_ vs retention time vs intensity) and can be obtained with the method `XICs`;
* __peak__ corresponds to a cluster of traces (e.g., integrated chromatographic MS peak) within an analysis file and can be accessed with the method `peaks`; 
* __feature__ corresponds to a group of corresponding peaks across analysis files/analyses and can be accessed with the method `features`;
* __MS2__ corresponds to second order fragmentation data as acquired in tandem MS or MS/MS mode and can be accessed with the method `MS2s`.

(...) `r # TODO add chromatograms entry and make method`

@@

```{r newstreamproject-call, message=FALSE, results='hide'}
sp1 <- newStreamProject(files = files[1:3], title = "Example_01")
```

@@

```{r msdata_streamproject_classes}
is(sp1)
```

The _msData_ structure includes the _streamProject_ class plus the `features` slot. The project, date and path slots correspond to the project title, creation date and directory in disk, respectively. The analyses slot is a list with a _msAnalysis_ S4 class object per MS file added. The structure of the _msAnalysis_ class is described in the next sub-chapter. The `features` slot is a _msFeatures_ S4 class object to hold the results from alignment, grouping and annotation of peaks across analyses. Further information for _msAnalysis_ is given in the next sub-chapter (\@ref(msanalysis-class))  and for _msFeatures_ is given in the sub-chapter \@ref(msfeatures-class), after creation of features by performing alignment and grouping of peaks across analyses.

@@

```{r structure_msdata}
str(sp1, max.level = 2)
```

@@

The _msAnalysis_ class is structured as shown below. Note that the S4 method `getAnalyses` for _msData_ was used to obtain the _msAnalysis_ of the first analysis as defined by the second argument. In general, access and plotting S4 methods can also be applied directly to the _msAnalysis_ class besides _msData_ (as described in \@ref(raw-data-access)) but is not part of the basic workflow and mostly used for advanced data visualization and access of low level data (e.g., traces).

@@

```{r msanalysis_structure}
a1 <- getAnalyses(sp1, 1)
str(a1, max.level = 2)
```

### Create msAnalysis

`r # TODO make method to open a file as xAnalysis``

### Handle metadata

@@

```{r eics-method}
#EIC of targets from first analysis using msAnalysis class
eic1 <- EICs(a1, mz = targets4)
head(eic1, 6)

@@

```{r ploteics-static}
#plot the produced EIC data.table
plotEICs(eic1)
```

```{r ploteics-interactive}
#using the msData class object and targets but with interactive plot
plotEICs(sp1, analyses = NULL, mz = targets4, colorBy = "analyses", interactive = TRUE)

@@

```{r tics_method}
# TIC for both analyses in the msData class object
tic1 <- TICs(sp1)
tic1

@@

```{r plotticss-method}
#using a data.table as obatined by TICs
plotTICs(tic1, interactive = TRUE)

@@

```{r xics-method}
#using the msData class object
xic1 <- XICs(sp1, mz = targets4)
xic1

@@

```{r plotxics-method}
#using the output of XICs S4 method (a data.table)
plotXICs(xic1, legendNames = c("target number 1", "target number 2"), plotTargetMark = TRUE,
 targetsMark = targets4[, c("mz", "rt")], ppmMark = 5, secMark = 10, numberRows = 2)

@@

```{r loadrawdata}
#load raw data to an msAnalysis object
a1 <- loadRawData(a1)

@@

```{r newstreamproject-basicworkflow, message=FALSE, results='hide'}
sp2 <- newStreamProject(files = files[c(10:12, 16:18)], path = getwd(), title = "Project Example")
```

## Manage analyses

### Add analyses

### Remove analyses

@@

```{r assign-replicates}
#getter for replicate names in msData
replicates(sp2)

@@

```{r assign-blanks}
#getter for blank replicates in each sample in a msData object
blanks(sp2)

@@

```{r other-s4methods}
#getter for analysis names in msdata
analyses(sp2)

@@

```{r create-processing-parameters}
sp2_pp <- sp2

@@

```{r peakpicking-msdata-run, results='hide', message=FALSE}
#peak picking call using the stored parameters
sp2_pp <- peakPicking(sp2_pp)
```

```{r peakpicking-msdata-show}
# peaks added for each analysis, as shown by column peaks of show method (below)
sp2_pp
```

```{r peakpicking-msdata-run2, results='hide', message=FALSE}
#alternatively, parameters can be added during the function call
sp2_pp_2 <- peakPicking(sp2, settings = settings_pp)
```

```{r peakpicking-msdata-show2, }
# parameters are added to each analysis after running peak picking
summary(getParameters(sp2_pp_2, where = "analyses", analyses = 1))

sp2_pp_2
```

```{r peakpicking-msanalysis-run, results='hide', message=FALSE}
# running peak picking for an msAnalysis object (settings were already added above)
a2 <- peakPicking(a2)
```

```{r peakpicking-msanalysis-show}
# msAnalysis with peaks
a2
```

@@

```{r peakpicking-multiple-settings}
sp2_pp_3 <- sp2

@@

# the algorithm of the influent samples is different
sapply(analyses(sp2_pp_3), function(x) getAlgorithm(getParameters(sp2_pp_3, where = "analyses", analyses = x)[[1]]))
```

```{r peakpicking-multiple-run, results='hide', message=FALSE}
#running peak picking
sp2_pp_3 <- peakPicking(sp2_pp_3)
```

```{r peakpicking-multiple-show}
# as the peak parameters with "openms" as algorithm were less stringent, the influent samples have more peaks then when "xcms3" is used 
sp2_pp_3
```

@@

Export settings:

```{r export-settings}

@@

#export settings as JSON
exportSettings(settings_pp_2, name = "settings_pp_2", format = "json", path = getwd())
```

Import settings:

```{r import-settings}
#import settings as rds
settings_pp_imported <- importSettings(file = paste0(getwd(), "/settings_pp.rds"))
all.equal(settings_pp, settings_pp_imported)

@@

```{r inspecting-peaks}
#access to peaks in msData based on built targets as used for extracting EICs
peaks(sp2_pp[c(1, 4)], mz = mz1[1], rt = rt1[1], ppm = 10, sec = 30)

@@

```{r creating-grouping-parameters}
param_g <- list(
  rtalign = TRUE,
  loadRawData = TRUE,
  groupParam = xcms::PeakDensityParam(
    sampleGroups = "holder",
    bw = 3,
    minFraction = 0.6,
    minSamples = 2,
    binSize = 0.008,
    maxFeatures = 100),
  preGroupParam = xcms::PeakDensityParam(
    sampleGroups = "holder",
    bw = 5,
    minFraction = 1,
    minSamples = 3,
    binSize = 0.008,
    maxFeatures = 100),
  retAlignParam = xcms::PeakGroupsParam(
    minFraction = 1,
    extraPeaks = 0,
    smooth = "loess",
    span = 0.3,
    family = "gaussian")
)

@@

```{r processing-grouping-run, results='hide', message=FALSE}
#using the added parameter settings
sp2_pg <- peakGrouping(sp2_pp)
```

```{r processing-grouping-show}
#features added to msData
sp2_pg
```

```{r processing_grouping-run2, results='hide', message=FALSE}
#adding the parameter settings as arguments
sp2_pg_2 <- peakGrouping(sp2_pp, settings = settings_pg)
```

```{r processing-grouping-show2}
#features added to msData
sp2_pg_2
```

@@

```{r msfeatures-class}
#accessing directly the msFeatures slot, not recommended
str(sp2_pg@features, max.level = 2)
```

@@

```{r inspecting-features}
#getter for feature intensities and intensity deviations
features(sp2_pg, mz = targets4)

@@

```{r plot-alignment}
plotAlignment(sp2_pg)
```

@@

```{r recursive-integration-settings}
#default settings S4 class with ChromPeakAreaParam from xcms as parameter settings
settings_fl <- fillingSettingsDefaultXCMS()

@@

```{r recursive-integration-run, results='hide', message=FALSE}
#filling peaks for msData
sp2_pg_fl <- peakFilling(sp2_pg, settings = settings_fl)
```

@@

```{r recursive-integration-inspection}
# inspecting the filled peaks for a given mz and rt target
plotFeaturePeaks(sp2_pg_fl, mz = 441.1670, rt = 916, ppm = 20, sec = 60)
```

@@

```{r creating-annotation-parameters, results='hide', message=FALSE}
#creating settings for using CAMERA
settings_an <- createSettings(
  call = "peakAnnotation",
  algorithm = "camera",
  settings = list(
    ionization = NA_character_,
    onlyIsotopes = FALSE,
    minSize = 1,
    relMinReplicates = 1,
    extraOpts = list(
      sigma = 6,
      perfwhm = 0.35,
      cor_eic_th = 0.3,
      graphMethod = "hcs",
      pval = 0.05,
      calcCiS = TRUE,
      calcIso = TRUE,
      calcCaS = TRUE,
      maxcharge = 3,
      maxiso = 5,
      ppm = 15,
      mzabs = 0.008,
      rules =  data.table::fread(system.file("rules/primary_adducts_pos.csv", package = "CAMERA"), header = TRUE),
      multiplier = 3,
      max_peaks = 500,
      intval = "maxo"
    )
  )
)
```

```{r annotation-run, results='hide', message=FALSE}
sp2_pg_fl_an <- peakAnnotation(sp2_pg_fl, settings = settings_an)
```

@@

```{r annotation-results}
#isotopic and adduct features annotated with a given target mz and rt
streamFind::annotation(sp2_pg_fl_an, mz = targets4[1, ], ppm = 20, rt = NULL, sec = 60, all = FALSE)

#plotting the annotation of each feature in a given target
plotAnnotation(sp2_pg_fl_an, mz = targets4[1, ], ppm = 20, rt = NULL, sec = 60, all = FALSE)
```

@@

```{r filter-subsetting}
#sub-setting the msData class on analyses (the 3 first)
sp2_pg[1:3, ]

#sub-setting the msData class on analys2s and features (the 3 last analyses and features annotated with a target)
target_s <- streamFind::annotation(sp2_pg_fl_an, mz = targets4[1, ], ppm = 20, rt = NULL, sec = 5, all = FALSE)
sp2_pg_fl_an[4:6, target_s$id]

@@

```{r filter-features}

@@

```{r filter-peaks}

@@

```{r quality-sn}

@@

```{r create-settings-msn}
#the parameter settings to obtained and process MSn spectra
param_msn <- list(
  isolationTimeWindow = 0,
  isolationMassWindow = 1.3,
  clusteringMethod = "distance",
  clusteringUnit = "ppm",
  clusteringWindow = 10,
  minIntensityPre = 200,
  minIntensityPost = 300,
  asPatRoon = FALSE, #when TRUE, is only used by running generateMS2
  mergeVoltages = TRUE,
  mergeBy = "analyses"
)

@@

```{r extract-ms2}
#head of MS2 data for target 1 in a msAnalysis object
head(MS2s(a1, mz = targets4[1, ], settings = settings_msn))

@@

## patRoon interchangeability

@@

```{r democode, echo=FALSE, eval=FALSE, include=FALSE}
